---
output: html_document
---
<h3> <center> Appendix </center> </h3>
  
Below are the libraries and data used for this analysis. 

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(broom)
#install.packages("table1")
library(table1)
library(htmlTable)

# Loading in the cleaned survey Data
survey_data <- read_csv("./outputs/survey_data.csv")
Newfoundland_survey_data <- read_csv("./outputs/survey_data_Newfoundland.csv")
PEY_survey_data <- read_csv("./outputs/survey_data_PEY.csv")
NovaScotia_survey_data <- read_csv("./outputs/survey_data_NovaScotia.csv")
NewBrunswick_survey_data <- read_csv("./outputs/survey_data_NewBrunswick.csv")
Quebec_survey_data <- read_csv("./outputs/survey_data_Quebec.csv")
Ontario_survey_data <- read_csv("./outputs/survey_data_Ontario.csv")
Manitoba_survey_data <- read_csv("./outputs/survey_data_Manitoba.csv")
Sas_survey_data <- read_csv("./outputs/survey_data_Sas.csv")
Alberta_survey_data <- read_csv("./outputs/survey_data_Alberta.csv")
BC_survey_data <- read_csv("./outputs/survey_data_BC.csv")
NorthernCanada_survey_data <- read_csv("./outputs/survey_data_NorthernCanada.csv")

# Loading in the cleaned census Data
census_data <- read_csv("./outputs/census_data.csv")
Newfoundland_census_data <- read_csv("./outputs/census_data_Newfoundland.csv")
PEY_census_data <- read_csv("./outputs/census_data_PEY.csv")
NovaScotia_census_data <- read_csv("./outputs/census_data_NovaScotia.csv")
NewBrunswick_census_data <- read_csv("./outputs/census_data_NewBrunswick.csv")
Quebec_census_data <- read_csv("./outputs/census_data_Quebec.csv")
Ontario_census_data <- read_csv("./outputs/census_data_Ontario.csv")
Manitoba_census_data <- read_csv("./outputs/census_data_Manitoba.csv")
Sas_census_data <- read_csv("./outputs/census_data_Sas.csv")
Alberta_census_data <- read_csv("./outputs/census_data_Alberta.csv")
BC_census_data <- read_csv("./outputs/census_data_BC.csv")
NorthernCanada_census_data <- read_csv("./outputs/census_data_NorthernCanada.csv")
raw_census_data <- read_csv("./outputs/not_grouped.csv")
```

Below are the logistic regression models created for each of the 11 regions (province/territory). 
```{r, message=FALSE, warning=FALSE}
model1 <- glm(cps19_vote_conservative ~ as.factor(is_working) + 
             as.factor(gender_sex) + 
               as.factor(is_Aboriginal) + as.factor(speak_French) + age_group + 
               as.factor(bornin_canada) + 
               household_income + education + as.factor(marital_status), 
           data=Newfoundland_survey_data, family="binomial")

model2 <- glm(cps19_vote_conservative ~ as.factor(is_working) + 
             as.factor(gender_sex) + 
               as.factor(is_Aboriginal) + as.factor(speak_French) + age_group + 
               as.factor(bornin_canada) + 
               household_income + education + as.factor(marital_status), 
           data=PEY_survey_data, family="binomial")

model3 <- glm(cps19_vote_conservative ~ as.factor(is_working) + 
             as.factor(gender_sex) +  
               as.factor(is_Aboriginal) + as.factor(speak_French) + age_group + 
               as.factor(bornin_canada) + 
               household_income + education + as.factor(marital_status), 
           data=NovaScotia_survey_data, family="binomial")

model4 <- glm(cps19_vote_conservative ~ as.factor(is_working) + 
             as.factor(gender_sex) +  
               as.factor(is_Aboriginal) + as.factor(speak_French) + age_group + 
               as.factor(bornin_canada) + 
               household_income + education + as.factor(marital_status), 
           data=NewBrunswick_survey_data, family="binomial")

model5 <- glm(cps19_vote_conservative ~ as.factor(is_working) + 
             as.factor(gender_sex) + 
               as.factor(is_Aboriginal) + as.factor(speak_French) + age_group + 
               as.factor(bornin_canada) + 
               household_income + education + as.factor(marital_status), 
           data=Quebec_survey_data, family="binomial")

model6 <- glm(cps19_vote_conservative ~ as.factor(is_working) + 
             as.factor(gender_sex) +
               as.factor(is_Aboriginal) + as.factor(speak_French) + age_group + 
               as.factor(bornin_canada) + 
               household_income + education + as.factor(marital_status), 
           data=Ontario_survey_data, family="binomial")

model7 <- glm(cps19_vote_conservative ~ as.factor(is_working) + 
             as.factor(gender_sex) +  
               as.factor(is_Aboriginal) + as.factor(speak_French) + age_group + 
               as.factor(bornin_canada) + 
               household_income + education + as.factor(marital_status), 
           data=Manitoba_survey_data, family="binomial")

model8 <- glm(cps19_vote_conservative ~ as.factor(is_working) + 
             as.factor(gender_sex) +  
               as.factor(is_Aboriginal) + as.factor(speak_French) + age_group + 
               as.factor(bornin_canada) + 
               household_income + education + as.factor(marital_status), 
           data=Sas_survey_data, family="binomial")

model9 <- glm(cps19_vote_conservative ~ as.factor(is_working) + 
             as.factor(gender_sex) +  
               as.factor(is_Aboriginal) + as.factor(speak_French) + age_group + 
               as.factor(bornin_canada) + 
               household_income + education + as.factor(marital_status), 
           data=Alberta_survey_data, family="binomial")

model10 <- glm(cps19_vote_conservative ~ as.factor(is_working) + 
             as.factor(gender_sex) + 
               as.factor(is_Aboriginal) + as.factor(speak_French) + age_group + 
               as.factor(bornin_canada) + 
               household_income + education + as.factor(marital_status), 
           data=BC_survey_data, family="binomial")

model11 <- glm(cps19_vote_conservative ~ as.factor(is_working) + 
             as.factor(gender_sex) + 
               as.factor(is_Aboriginal) + as.factor(speak_French) + age_group + 
               as.factor(bornin_canada) + 
               household_income + education + as.factor(marital_status), 
           data=NorthernCanada_survey_data, family="binomial")
```

Below are tables with summary data of the logistic regression model used to perform analysis on a provincial basis. Tables A to K include the slope estimate, standard error, test statistic and p-values of the variables in the model for each province/territory.  
```{r, echo=FALSE}
# Newfoundland and the Labrador
temp1 <- broom::tidy(model1)

temp1 <- select(temp1,-c(term))

temp1 %>% mutate_if(is.numeric, round, 4) %>% 
  addHtmlTableStyle(align = "r", 
                    css.cell=c("padding-left: 15px; padding-right:15px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:15px;"),
                    css.header="font-weight: normal;") %>% 
  htmlTable(rnames = c("Intercept", "Is Employed", "Is Female", "Is Aboriginal",
                       "Speaks French", "Age Group", "Born In Canada", 
                       "Household Income", "Education Level", "Legally Married",
                       "Living Common Law", "Separated", "Divorced", "Widowed"),
            header = c("Slope Estimate", "Standard Error", "Test Statistic",
                       "P-Values"),
            caption = "Table A: Summary of Logistic Regression Model for the Newfoundland and the Labrador Province")
```

```{r, echo=FALSE}
# Prince Edward Island
temp2 <- broom::tidy(model2)

temp2 <- select(temp2,-c(term))

temp2 %>% mutate_if(is.numeric, round, 4) %>% 
  addHtmlTableStyle(align = "r", 
                    css.cell=c("padding-left: 15px; padding-right:15px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:15px;"),
                    css.header="font-weight: normal;") %>% 
  htmlTable(rnames = c("Intercept", "Is Employed", "Is Female", "Is Aboriginal",
                       "Speaks French", "Age Group", "Born In Canada", 
                       "Household Income", "Education Level", "Legally Married",
                       "Living Common Law", "Separated", "Divorced", "Widowed"),
            header = c("Slope Estimate", "Standard Error", "Test Statistic",
                       "P-Values"),
            caption = "Table B: Summary of Logistic Regression Model for Prince Edward Island")
```

```{r, echo=FALSE}
# Nova Scotia
temp3 <- broom::tidy(model3)

temp3 <- select(temp3,-c(term))

temp3 %>% mutate_if(is.numeric, round, 4) %>% 
  addHtmlTableStyle(align = "r", 
                    css.cell=c("padding-left: 15px; padding-right:15px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:15px;"),
                    css.header="font-weight: normal;") %>% 
  htmlTable(rnames = c("Intercept", "Is Employed", "Is Female", "Is Aboriginal",
                       "Speaks French", "Age Group", "Born In Canada", 
                       "Household Income", "Education Level", "Legally Married",
                       "Living Common Law", "Separated", "Divorced", "Widowed"),
            header = c("Slope Estimate", "Standard Error", "Test Statistic",
                       "P-Values"),
            caption = "Table C: Summary of Logistic Regression Model for the Province of Nova Scotia")
```

```{r, echo=FALSE}
# New Brunswick
temp4 <- broom::tidy(model4)

temp4 <- select(temp4,-c(term))

temp4 %>% mutate_if(is.numeric, round, 4) %>% 
  addHtmlTableStyle(align = "r", 
                    css.cell=c("padding-left: 15px; padding-right:15px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:15px;"),
                    css.header="font-weight: normal;") %>% 
  htmlTable(rnames = c("Intercept", "Is Employed", "Is Female", "Is Aboriginal",
                       "Speaks French", "Age Group", "Born In Canada", 
                       "Household Income", "Education Level", "Legally Married",
                       "Living Common Law", "Separated", "Divorced", "Widowed"),
            header = c("Slope Estimate", "Standard Error", "Test Statistic",
                       "P-Values"),
            caption = "Table D: Summary of Logistic Regression Model for the Province of New Brunswick")
```

```{r, echo=FALSE}
# Quebec
temp5 <- broom::tidy(model5)

temp5 <- select(temp5,-c(term))

temp5 %>% mutate_if(is.numeric, round, 4) %>% 
  addHtmlTableStyle(align = "r", 
                    css.cell=c("padding-left: 15px; padding-right:15px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:15px;"),
                    css.header="font-weight: normal;") %>% 
  htmlTable(rnames = c("Intercept", "Is Employed", "Is Female", "Is Aboriginal",
                       "Speaks French", "Age Group", "Born In Canada", 
                       "Household Income", "Education Level", "Legally Married",
                       "Living Common Law", "Separated", "Divorced", "Widowed"),
            header = c("Slope Estimate", "Standard Error", "Test Statistic",
                       "P-Values"),
            caption = "Table E: Summary of Logistic Regression Model for Quebec")
```

```{r, echo=FALSE}
# Ontario
temp6 <- broom::tidy(model6)

temp6 <- select(temp6,-c(term))

temp6 %>% mutate_if(is.numeric, round, 4) %>% 
  addHtmlTableStyle(align = "r", 
                    css.cell=c("padding-left: 15px; padding-right:15px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:15px;"),
                    css.header="font-weight: normal;") %>% 
  htmlTable(rnames = c("Intercept", "Is Employed", "Is Female", "Is Aboriginal",
                       "Speaks French", "Age Group", "Born In Canada", 
                       "Household Income", "Education Level", "Legally Married",
                       "Living Common Law", "Separated", "Divorced", "Widowed"),
            header = c("Slope Estimate", "Standard Error", "Test Statistic",
                       "P-Values"),
            caption = "Table F: Summary of Logistic Regression Model for Ontario")
```

```{r, echo=FALSE}
# Manitoba
temp7 <- broom::tidy(model7)

temp7 <- select(temp7,-c(term))

temp7 %>% mutate_if(is.numeric, round, 4) %>% 
  addHtmlTableStyle(align = "r", 
                    css.cell=c("padding-left: 15px; padding-right:15px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:15px;"),
                    css.header="font-weight: normal;") %>% 
  htmlTable(rnames = c("Intercept", "Is Employed", "Is Female", "Is Aboriginal",
                       "Speaks French", "Age Group", "Born In Canada", 
                       "Household Income", "Education Level", "Legally Married",
                       "Living Common Law", "Separated", "Divorced", "Widowed"),
            header = c("Slope Estimate", "Standard Error", "Test Statistic",
                       "P-Values"),
            caption = "Table G: Summary of Logistic Regression Model for Manitoba")
```

```{r, echo=FALSE}
# Saskatchewan
temp8 <- broom::tidy(model8)

temp8 <- select(temp8,-c(term))

temp8 %>% mutate_if(is.numeric, round, 4) %>% 
  addHtmlTableStyle(align = "r", 
                    css.cell=c("padding-left: 15px; padding-right:15px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:15px;"),
                    css.header="font-weight: normal;") %>% 
  htmlTable(rnames = c("Intercept", "Is Employed", "Is Female", "Is Aboriginal",
                       "Speaks French", "Age Group", "Born In Canada", 
                       "Household Income", "Education Level", "Legally Married",
                       "Living Common Law", "Separated", "Divorced", "Widowed"),
            header = c("Slope Estimate", "Standard Error", "Test Statistic",
                       "P-Values"),
            caption = "Table H: Summary of Logistic Regression Model for Saskatchewan")
```

```{r, echo=FALSE}
# Alberta
temp9 <- broom::tidy(model9)

temp9 <- select(temp9,-c(term))

temp9 %>% mutate_if(is.numeric, round, 4) %>% 
  addHtmlTableStyle(align = "r", 
                    css.cell=c("padding-left: 15px; padding-right:15px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:15px;"),
                    css.header="font-weight: normal;") %>% 
  htmlTable(rnames = c("Intercept", "Is Employed", "Is Female", "Is Aboriginal",
                       "Speaks French", "Age Group", "Born In Canada", 
                       "Household Income", "Education Level", "Legally Married",
                       "Living Common Law", "Separated", "Divorced", "Widowed"),
            header = c("Slope Estimate", "Standard Error", "Test Statistic",
                       "P-Values"),
            caption = "Table I: Summary of Logistic Regression Model for Alberta")
```

```{r, echo=FALSE}
# British Columbia
temp10 <- broom::tidy(model10)

temp10 <- select(temp10,-c(term))

temp10 %>% mutate_if(is.numeric, round, 4) %>% 
  addHtmlTableStyle(align = "r", 
                    css.cell=c("padding-left: 15px; padding-right:15px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:15px;"),
                    css.header="font-weight: normal;") %>% 
  htmlTable(rnames = c("Intercept", "Is Employed", "Is Female", "Is Aboriginal",
                       "Speaks French", "Age Group", "Born In Canada", 
                       "Household Income", "Education Level", "Legally Married",
                       "Living Common Law", "Separated", "Divorced", "Widowed"),
            header = c("Slope Estimate", "Standard Error", "Test Statistic",
                       "P-Values"),
            caption = "Table J: Summary of Logistic Regression Model for British Columbia")
```

```{r, echo=FALSE}
# Northern Canada (Nunavut, Yukon, Northwest Territories)
temp11 <- broom::tidy(model11)

temp11 <- select(temp11,-c(term))

temp11 %>% mutate_if(is.numeric, round, 4) %>% 
  addHtmlTableStyle(align = "r", 
                    css.cell=c("padding-left: 15px; padding-right:15px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:30px;", 
                               "padding-left: 30px; padding-right:15px;"),
                    css.header="font-weight: normal;") %>% 
  htmlTable(rnames = c("Intercept", "Is Employed", "Is Female", "Is Aboriginal",
                       "Speaks French", "Age Group", "Born In Canada", 
                       "Household Income", "Education Level", "Legally Married",
                       "Living Common Law", "Separated", "Divorced", "Widowed"),
            header = c("Slope Estimate", "Standard Error", "Test Statistic",
                       "P-Values"),
            caption = "Table K: Summary of Logistic Regression Model for Northern Canada")
```


